---
title: "blog_6"
author: "Carine Hajjar"
date: "10/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(statebins)
library(stargazer)
library(cowplot)
library(gt)

```

# Data 
```{r}
demog <- read_csv("~/Desktop/R studio/carine-h.github.io/data/demographic_1990-2018.csv")
pvstate_df    <- read_csv("~/Desktop/R studio/carine-h.github.io/data/popvote_bystate_1948-2016.csv")
pollstate_df  <- read_csv("~/Desktop/R studio/carine-h.github.io/data/pollavg_bystate_1968-2016.csv")
hispanic_2020 <- read_csv("~/Desktop/R studio/carine-h.github.io/data/csvData.csv")
race_2020 <- read_csv("~/Desktop/R studio/carine-h.github.io/data/demog_race.csv")
electoral_votes <- read_csv("~/Desktop/R studio/carine-h.github.io/data/electoralcollegevotes_1948-2020.csv")

# tidyinig electoral votes for my purposes 
electoral <- electoral_votes %>%
  select(X1, '2020') %>%
  rename(state = X1) %>%
  mutate(state = state.abb[match(state,state.name)]) %>%
  na.omit() 

# each data set 
demog # age, race, state, year, and count
pvstate_df # state by state popular vote 
pollstate_df # poll numbers 

# state names and abbreviations
pvstate_df$state <- state.abb[match(pvstate_df$state, state.name)]
pollstate_df$state <- state.abb[match(pollstate_df$state, state.name)]

dat <- pvstate_df %>% 
  full_join(pollstate_df %>% 
              filter(weeks_left == 3) %>% 
              group_by(year,party,state) %>% 
              summarise(avg_poll=mean(avg_poll)),
            by = c("year" ,"state")) %>%
  left_join(demog %>%
              select(-c("total")),
            by = c("year" ,"state"))
# demographics, poll numbers, and popular vote 

dat$region <- state.division[match(dat$state, state.abb)]
demog$region <- state.division[match(demog$state, state.abb)]

dat_change <- dat %>%
  group_by(state) %>%
  mutate(Asian_change = Asian - lag(Asian, order_by = year),
         Black_change = Black - lag(Black, order_by = year),
         Hispanic_change = Hispanic - lag(Hispanic, order_by = year),
         Indigenous_change = Indigenous - lag(Indigenous, order_by = year),
         White_change = White - lag(White, order_by = year),
         Female_change = Female - lag(Female, order_by = year),
         Male_change = Male - lag(Male, order_by = year),
         age20_change = age20 - lag(age20, order_by = year),
         age3045_change = age3045 - lag(age3045, order_by = year),
         age4565_change = age4565 - lag(age4565, order_by = year),
         age65_change = age65 - lag(age65, order_by = year)
  )

# new data for 2020
demog_2020 <- subset(demog, year == 2018)
demog_2020 <- as.data.frame(demog_2020)
rownames(demog_2020) <- demog_2020$state
demog_2020 <- demog_2020[state.abb, ] # prints regions of the US 

# calculating demographic changes from 2016-2018
demog_2020_change <- demog %>%
  filter(year %in% c(2016, 2018)) %>%
  group_by(state) %>%
  mutate(Asian_change = Asian - lag(Asian, order_by = year), # CALCULATING CHANGES IN POPULATION
         Black_change = Black - lag(Black, order_by = year),
         Hispanic_change = Hispanic - lag(Hispanic, order_by = year),
         Indigenous_change = Indigenous - lag(Indigenous, order_by = year),
         White_change = White - lag(White, order_by = year),
         Female_change = Female - lag(Female, order_by = year),
         Male_change = Male - lag(Male, order_by = year),
         age20_change = age20 - lag(age20, order_by = year),
         age3045_change = age3045 - lag(age3045, order_by = year),
         age4565_change = age4565 - lag(age4565, order_by = year),
         age65_change = age65 - lag(age65, order_by = year)
  ) %>%
  filter(year == 2018)

demog_2020_change <- as.data.frame(demog_2020_change)
rownames(demog_2020_change) <- demog_2020_change$state
demog_2020_change <- demog_2020_change[state.abb, ]

```

# Model
```{r}
#####------------------------------------------------------#
#####  Proposed models ####
#####------------------------------------------------------#

mod_demog_change <- lm(D_pv2p ~ Black_change + Hispanic_change + Asian_change +
                         Female_change +
                         age3045_change + age4565_change + age65_change +
                         as.factor(region), data = dat_change)
# democratic vote share is positively correlated with Asian, Black, Female, Hispanic, 30-45
```


# My 2020 Realistic Forecast 
I want to make a more realistic projection by using *current* demographic information
```{r, fig.width= 10}
# updated demographics
dat_2020 <- race_2020 %>%
  left_join(hispanic_2020, by = "State") %>%
  mutate(Hispanic = 100*HispanicPerc, 
         state = State, 
         White = 100*WhitePerc, 
         Asian = 100*AsianPerc, 
         Black = 100*BlackPerc) %>%
  select(state, Hispanic, White, Asian, Black) %>%
  mutate(state = state.abb[match(state,state.name)]) %>%
  na.omit() %>%
  mutate(year = 2020)

# 2018 demographics
dat_2018 <- demog %>%
  filter(year == 2018) 

# joining demographics
real_2020_change <- bind_rows(dat_2018, dat_2020)


# calculating percent changes in available demographic groups
## I used 0 percent change with populations that lacked demographic data (age and gender)
real_2020 <- real_2020_change %>%
  filter(year %in% c(2018, 2020)) %>%
  group_by(state) %>%
  mutate(Asian_change = Asian - lag(Asian, order_by = year), # CALCULATING CHANGES IN POPULATION
         Black_change = Black - lag(Black, order_by = year),
         Hispanic_change = Hispanic - lag(Hispanic, order_by = year),
         Indigenous_change = 0,
         White_change = White - lag(White, order_by = year),
         Female_change = 0,
         Male_change = 0,
         age20_change = 0,
         age3045_change = 0,
         age4565_change = 0,
         age65_change = 0) %>%
  filter(year == 2020)

real_2020 <- as.data.frame(real_2020)
rownames(real_2020) <- real_2020$state
real_2020 <- real_2020[state.abb, ]
real_2020$region <- state.division[match(real_2020$state, state.abb)]

# prediction using actual 2020 data
predict(mod_demog_change, newdata = real_2020) 

his_original <- tibble(predict(mod_demog_change, newdata = demog_2020_change), state = state.abb, pred = `predict(mod_demog_change, newdata = demog_2020_change)`)
his_2020 <- tibble(predict(mod_demog_change, newdata = real_2020), state = state.abb, pred = `predict(mod_demog_change, newdata = real_2020)`)

plot_original <- his_original %>%  ##`statebins` needs state to be character, not factor!
  mutate(state = as.character(state)) %>%
  ggplot(aes(state = state, fill = (pred >= 50))) +
  geom_statebins() +
  theme_statebins() +
  labs(title = "Original 2020 Presidential Election Prediction",
       subtitle = "Demographic Change Effect \n Using 2016-2018 Demographic Shifts",
       fill = "") +
  theme(legend.position = "none", 
        plot.subtitle = element_text(size = 7, hjust = 0.5),
        plot.title   = element_text(size = 9, hjust = 0.5, face = "bold"))

plot_1 <- his_2020 %>% 
  mutate(state = as.character(state)) %>% ##`statebins` needs state to be character, not factor!
  ggplot(aes(state = state, fill = (pred >= 50))) +
  geom_statebins() +
  theme_statebins() +
  labs(title = "Updated 2020 Presidential Election Prediction",
       subtitle = "Demographic Change Effect \n Using 2018-2020 Demographic Shifts",
       fill = "", 
       caption = "Data: World Population Review") +
  theme(legend.position = "none", 
        plot.subtitle = element_text(size = 7, hjust = 0.5),
        plot.title   = element_text(size = 9, hjust = 0.5, face = "bold"))

plot_2020 <- plot_grid(plot_original, plot_1)

ggsave("~/Desktop/R studio/carine-h.github.io/images/2020_plot.png", plot_2020)
```

```{r}
# 2020 PREDICTION 
# original prediction counting electoral 
his_original %>%
  select(state, pred) %>%
  left_join(electoral, by = "state") %>%
  mutate(state_win = case_when(pred > 50 ~ "Biden",
                            pred < 50 ~ "Trump")) %>%
  group_by(state_win) %>%
  summarise(electoral_votes = sum(`2020`)) %>%
  gt() %>%
  tab_header(title = md("**2020 Electoral Outcome Using 2016-2018 Demographics**"), 
               subtitle = "Trump Wins") %>%
  cols_label(state_win = md("**Candidate**"),
               electoral_votes = md("**Total Electoral Votes**")) 
# gtsave(orig, "tab_original.html", path = "~/Desktop/R studio/carine-h.github.io/images")
  
# 2020 updated demographics counting electorals
his_2020 %>%
  select(state, pred) %>%
  left_join(electoral, by = "state") %>%
  mutate(state_win = case_when(pred > 50 ~ "Biden",
                            pred < 50 ~ "Trump")) %>%
  group_by(state_win) %>%
  summarise(electoral_votes = sum(`2020`)) %>%
  gt() %>%
  tab_header(title = md("**2020 Electoral Outcome Using 2020 Demographics**"), 
               subtitle = "Biden Wins") %>%
   cols_label(state_win = md("**Candidate**"),
               electoral_votes = md("**Total Electoral Votes**")) %>%
  tab_source_note(md("*Data: World Population Review*"))

```
- (https://worldpopulationreview.com/states/states-by-race)

# 2050 Demographic Forecasts 
- note: since there aren't state-by-state predictions for 2050, just national demographic predictions, I applied nation-wide changes on the state-by-state level
  - obviously, some states would not experience this level of shift but this is a useful exercise 
- Using 2050's predicted demographics: what would this election look like?
    - https://www.pewresearch.org/hispanic/2008/02/11/us-population-projections-2005-2050/
```{r}
#####------------------------------------------------------#
#####  Proposed models ####
#####------------------------------------------------------#

mod_demog_change <- lm(D_pv2p ~ Black_change + Hispanic_change + Asian_change +
                         Female_change +
                         age3045_change + age4565_change + age65_change +
                         as.factor(region), data = dat_change)
# democratic vote share is positively correlated with Asian, Black, Female, Hispanic, 30-45
```

```{r, fig.width = 10}
# Demographic change according to Pew: https://www.pewresearch.org/fact-tank/2019/04/11/6-demographic-trends-shaping-the-u-s-and-the-world-in-2019/
mean(demog_2020_change$Hispanic) # 10.79 hispanic in 2018 but in Pew it was 11.9 in 2016 and then rose to 13.3 (these are electorally eligible)
mean(demog_2020_change$White) # 71.62
mean(demog_2020_change$Black) # 10.62
mean(demog_2020_change$Asian) # 5.37
# Pew voter demog percent change estimate 2050: 
  # White - 47% 
  # Hispanic - 29%
  # Black - 13% 
  # Asian - 9%
# calculating percent change from 2018 values and multiplying coefficients to simulate population change 
  # Hispanic:
    his_change <- ((29-10.79)/10.79)*0.64
  # Black 
    black_change <- ((13-10.62)/13)*5.69
  # Asian
    asian_change <- ((9-5.37)/5.37)*0.19


# new data for 2020
demog_2020 <- subset(demog, year == 2018)
demog_2020 <- as.data.frame(demog_2020)
rownames(demog_2020) <- demog_2020$state
demog_2020 <- demog_2020[state.abb, ] # prints regions of the US 

# calculating demographic changes from 2016-2018
demog_2020_change <- demog %>%
  filter(year %in% c(2016, 2018)) %>%
  group_by(state) %>%
  mutate(Asian_change = Asian - lag(Asian, order_by = year), # CALCULATING CHANGES IN POPULATION
         Black_change = Black - lag(Black, order_by = year),
         Hispanic_change = Hispanic - lag(Hispanic, order_by = year),
         Indigenous_change = Indigenous - lag(Indigenous, order_by = year),
         White_change = White - lag(White, order_by = year),
         Female_change = Female - lag(Female, order_by = year),
         Male_change = Male - lag(Male, order_by = year),
         age20_change = age20 - lag(age20, order_by = year),
         age3045_change = age3045 - lag(age3045, order_by = year),
         age4565_change = age4565 - lag(age4565, order_by = year),
         age65_change = age65 - lag(age65, order_by = year)
  ) %>%
  filter(year == 2018)

demog_2020_change <- as.data.frame(demog_2020_change)
rownames(demog_2020_change) <- demog_2020_change$state
demog_2020_change <- demog_2020_change[state.abb, ]

# prediction
pred_2050 <- predict(mod_demog_change, newdata = demog_2020_change) + 
  his_change*demog_2020$Hispanic +
  black_change*demog_2020$Black +
  asian_change*demog_2020$Asian
# this is an 11.8% increase in voting by increasing coefficient by same percentage 

his_original <- tibble(predict(mod_demog_change, newdata = demog_2020_change), state = state.abb, pred = `predict(mod_demog_change, newdata = demog_2020_change)`)
his_2050 <- tibble(predict(mod_demog_change, newdata = demog_2020_change) + 
  his_change*demog_2020$Hispanic +
  black_change*demog_2020$Black +
  asian_change*demog_2020$Asian, state = state.abb, pred = `+...`)

plot_original <- his_original %>%  ##`statebins` needs state to be character, not factor!
  mutate(state = as.character(state)) %>%
  ggplot(aes(state = state, fill = (pred >= 50))) +
  geom_statebins() +
  theme_statebins() +
  labs(title = "2020 Presidential Election Prediction",
       subtitle = "Historical Demographic Change Effect",
       fill = "") +
  theme(legend.position = "none", 
        plot.subtitle = element_text(size = 7, hjust = 0.5),
        plot.title   = element_text(size = 9, hjust = 0.5, face = "bold"))

plot_1 <- his_2050 %>% 
  mutate(state = as.character(state)) %>% ##`statebins` needs state to be character, not factor!
  ggplot(aes(state = state, fill = (pred >= 50))) +
  geom_statebins() +
  theme_statebins() +
  labs(title = "2020 Presidential Election Prediction",
       subtitle = "2050 National Demographic Changes",
       fill = "") +
  theme(legend.position = "none", 
        plot.subtitle = element_text(size = 7, hjust = 0.5),
        plot.title   = element_text(size = 9, hjust = 0.5, face = "bold"))

plot_2050 <- plot_grid(plot_original, plot_1)

ggsave("~/Desktop/R studio/carine-h.github.io/images/2050_plot.png", plot_2050)
```

```{r}
# 2050 PREDICTION
his_2050 %>%
  select(state, pred) %>%
  left_join(electoral, by = "state") %>%
  mutate(state_win = case_when(pred > 50 ~ "Biden",
                            pred < 50 ~ "Trump")) %>%
  group_by(state_win) %>%
  summarise(electoral_votes = sum(`2020`)) %>%
  gt() %>%
  tab_header(title = md("**2020 Electoral Outcome Using 2050 National Demographics**"), 
               subtitle = "Biden Wins") %>%
   cols_label(state_win = md("**Candidate**"),
               electoral_votes = md("**Total Electoral Votes**")) %>%
  tab_source_note(md("*Data: Pew*")) 
```
- (https://www.pewresearch.org/fact-tank/2019/04/11/6-demographic-trends-shaping-the-u-s-and-the-world-in-2019/)

