---
title: "blog_narrative_study"
author: "Carine Hajjar"
date: "11/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(quanteda) ## package for analyzing text-as-data
library(ggplot2)
library(readxl)
library(usmap)
library(stringr)
library(tidyverse)
library(gridExtra)
library(ggpubr)
library(webshot)


# loading in data 
speech_df <- read_csv("~/Desktop/R studio/carine-h.github.io/data/campaignspeech_2019-2020.csv")
pv_by_county_2000_2016_10 <- read_csv("~/Desktop/R studio/carine-h.github.io/data/popvote_bycounty_2000-2016.csv")
pv_by_county_2020_10 <- read_csv("~/Desktop/R studio/carine-h.github.io/data/popvote_bycounty_2020.csv")
hisp_black_county_2018 <- read_xlsx("~/Desktop/R studio/carine-h.github.io/data/2018_hisp_black_counties.xlsx")
fb_ad_df <- read_csv("~/Desktop/R studio/carine-h.github.io/data/fb_ad.csv")
fb_location_df <- read_csv("~/Desktop/R studio/carine-h.github.io/data/fb_location.csv")


state <- pv_by_county_2000_2016_10 %>%
  select(county, state, fips)

# cleaning 2020 county data to match 2000-2016
  # will have to create a D vote margin
pv_by_county_2020_10 = pv_by_county_2020_10[-1,]
pv_2020 <- pv_by_county_2020_10 %>%
  select("FIPS", "Geographic Name", "Geographic Subtype", "Total Vote", "Joseph R. Biden Jr.", "Donald J. Trump") %>%
  rename(county = 'Geographic Name', 
         type = 'Geographic Subtype',
         total = 'Total Vote',
         biden = 'Joseph R. Biden Jr.', 
         trump = 'Donald J. Trump', 
         fips = 'FIPS') %>%
  mutate(total = as.numeric(total), 
         biden = as.numeric(biden), 
         trump = as.numeric(trump), 
         fips = as.numeric(fips)) %>%
  mutate(D_pv2p = biden/total, 
         R_pv2p = trump/total, 
         D_win_margin = 100* ((biden-trump)/total)) %>%
  select(county, D_win_margin, fips) %>%
  mutate(year = 2020) %>%
  left_join(state, by = c("fips", "county")) %>%
  unique()

# pop vote 2000-2020
pv_change <- pv_by_county_2000_2016_10 %>%
  select(fips, year, state, county, D_win_margin) %>%
  rbind(pv_2020) %>%
  filter(year >= 2016) %>%
  spread(year, D_win_margin) %>%
  rename(d_marg_2016 = '2016', 
         d_marge_2020 = '2020') %>%
  mutate(d_marg_change = d_marge_2020 - d_marg_2016) 

# hispanic and black vote
hb <- hisp_black_county_2018 %>%
  mutate(county = str_remove_all(county, "County")) %>%
  filter(!str_detect(county, "Danville city"), 
         !str_detect(county, "Petersburg city"),
         !str_detect(county, "Portsmouth city"),
         !str_detect(county, "Richmond city"))

# overall votes 2020 
pv_2020_votes <- pv_by_county_2020_10 %>%
  select("FIPS", "Geographic Name", "Geographic Subtype", "Total Vote", "Joseph R. Biden Jr.", "Donald J. Trump") %>%
  rename(county = 'Geographic Name', 
         type = 'Geographic Subtype',
         total = 'Total Vote',
         biden = 'Joseph R. Biden Jr.', 
         trump = 'Donald J. Trump', 
         fips = 'FIPS') %>%
  mutate(total = as.numeric(total), 
         biden = as.numeric(biden), 
         trump = as.numeric(trump), 
         fips = as.numeric(fips)) %>%
  mutate(D_pv2p = biden/total, 
         R_pv2p = trump/total, 
         D_win_margin = 100* ((biden-trump)/total)) %>%
  left_join(change.dat, by = c("county"))

# write.csv(pv_change, "~/Desktop/R studio/carine-h.github.io/data/pv_change.csv", row.names = FALSE)
# write.csv(hb, "~/Desktop/R studio/carine-h.github.io/data/hispanic_counties.csv", row.names = FALSE)

vote.dat = read_csv("~/Desktop/R studio/carine-h.github.io/data/pv_change.csv")
change.dat = read_csv("~/Desktop/R studio/carine-h.github.io/data/hispanic_counties.csv")


# HISPANIC COUNTIES CHANGE IN D VOTE MARGIN 
dat <-  left_join(change.dat,vote.dat) %>%
  filter(largest_racial_ethnic_group == "Hispanic") %>%
  left_join(pv_2020_votes) %>%
  select(state, county, percent_non_white, one_nonwhite_group, fips, d_marg_2016, d_marge_2020, d_marg_change, D_pv2p, R_pv2p, D_win_margin) %>%
  mutate(fips = sprintf("%05d", fips))
```

- narrative: Hispanics made this race a close call for Biden (at least one factor) and reaching this demographic was a huge failure for Democrats 


# Notes on Pew Data
[Pew](https://www.pewresearch.org/fact-tank/2019/11/20/in-a-rising-number-of-u-s-counties-hispanic-and-black-americans-are-the-majority/) info using Census 
- note from Pew on data: "This analysis includes only counties with 10,000 or more residents in 2018. These counties account for 77% of the nation’s 3,142 counties and 99% of the U.S. population."
- CITATION: Pew Research Center analysis of 2000 decennial census and 2018 Census Bureau population estimates.
- official name of table: "Population in U.S. counties where Hispanic, black, or indigenous people are a **large** share of residents"
  - column names: "percent of population that was one racial/ethnic group other than white in 2018" AND "largest racial/ethnic group, 2018"
- article: by Katherine Schaeffer
  - 2018: 151 US counties where Hispanics, blacks, and American Indians/Alaska Natives made up majority of pop
    - up from 110 in 2000 
    - 41 of new were black or Hispanic 

# Overall relationship between prop hispanic and d vote 2020 
```{r}
county_rel <- dat %>%
ggplot(aes(one_nonwhite_group, D_pv2p))+
  geom_point() +
  geom_smooth(method = lm) +
  geom_hline(yintercept = 0.5, 
             colour = "red", 
             linetype = "dashed")+
  labs(title = "The Relationship Between the Percentage of the \n Hispanic Population and the 2020 Democratic Vote Share", 
       subtitle = "Hispanic-Majority Counties", 
       x = "Percentage of Hispanic Citizens per County", 
       y= "Democratic Vote Share") +
  theme(plot.title = element_text(face = "bold")) 

ggsave("~/Desktop/R studio/carine-h.github.io/images/10_county_rel.png", county_rel)
```





# Plots
```{r}
#####-------------------------------------------------------------------------------#
##### Hist and Scatter ####
#####-------------------------------------------------------------------------------#

states <- c("Arizona", "California", "Florida", "Kansas", "New Mexico", "New York", "Texas", "Washington")

png("~/Desktop/R studio/carine-h.github.io/images/10_hist_d_change.png", width=800, height=400)
hist(dat$d_marg_change, 
     main = "Hispanic-Majority County-Level Change in \nDemocratic Vote Share Margin: 2016-2020", 
     col = "dodgerblue", 
     xlab = "Change in Democratic Margin of Victory")
abline(v = mean(dat$d_marg_change, na.rm = TRUE), col = 'red', lwd = 2)
dev.off()

pv_change %>%
  ggplot(aes(d_marge_2020, d_marg_2016))+ 
  geom_point()+
  geom_smooth(method = lm)

# all together
  # Q does this look good tho???
m1 <- plot_usmap(data = dat, regions = "counties", values = "d_marg_change", 
           include = states, 
           exclude = "New York") +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "Win Margin") +
  theme_void()+
  labs(title = "2016-2020 Change in Democratic Win Margins",
  subtitle = "Hispanic-Majority Counties") +
  theme(plot.title = element_text(face = "bold"))

ggsave("~/Desktop/R studio/carine-h.github.io/images/10_dem_comp.png", m1)

m2 <- plot_usmap(data = dat, regions = "counties", values = "D_pv2p", 
           include = states, 
           exclude = "New York") +
  scale_fill_stepsn(breaks=c(dat$D_pv2p,seq(0, .50, 1)),
                    colours=c("red", "white", "blue")) +
  theme_void()+
  labs(title = "2020 Winners in Hispanic-Majority Counties", 
       subtitle = "Presidential Election 2020") +
  theme(plot.title = element_text(face = "bold")) +
  theme(legend.position = "none")

ggsave("~/Desktop/R studio/carine-h.github.io/images/10_dem_win.png", m2)
    
#####-------------------------------------------------------------------------------#
##### USMAPS prep ####
#####-------------------------------------------------------------------------------#

# extracting just counties we want 
fip_fl <- dat$fips[dat$state == "Florida"]
fip_tx <- dat$fips[dat$state == "Texas"]
fip_nm <- dat$fips[dat$state == "New Mexico"]
fip_ks <- dat$fips[dat$state == "Kansas"]
fip_az <- dat$fips[dat$state == "Arizona"]
fip_cali <- dat$fips[dat$state == "California"]
```


```{r}
#####-------------------------------------------------------------------------------#
##### Florida map ####
#####-------------------------------------------------------------------------------#
# FLORIDA
fl1 <- plot_usmap(data = dat, regions = "counties", values = "d_marg_change", 
           include = c(fip_fl),
           labels = TRUE) +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "win margin") +
  theme_void()+
  theme(legend.position = "none") 

fl2 <- plot_usmap(data = dat, regions = "counties", values = "d_marg_change", 
           include = "Florida") +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "Win Margin") +
  theme_void() 


fl_comp <- ggarrange(fl1, fl2, ncol=2)
florida_comp <- annotate_figure(fl_comp,
               top = text_grob("Florida 2016-2020: Change in Democratic \n Win Margins in Hispanic-Majority Counties", face = "bold", size = 13), 
               bottom = text_grob("Source: Pew",
                                  hjust = 1, x = 0.85, face = "italic", size = 10))

ggsave("~/Desktop/R studio/carine-h.github.io/images/10_florida_comp.png", florida_comp)
```
annotate_figure(florida_comp,
               top = text_grob("Florida 2016-2020: Change in Democratic Win Margins in Hispani-Majority Counties", face = "bold", size = 14),
               bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               left = text_grob("Figure arranged using ggpubr", color = "green", rot = 90),
               right = text_grob(bquote("Superscript: ("*kg~NH[3]~ha^-1~yr^-1*")"), rot = 90),
               fig.lab = "Figure 1", fig.lab.face = "bold"
)
```{r}
#####-------------------------------------------------------------------------------#
##### Texas map ####
#####-------------------------------------------------------------------------------#
# TEXAS
p_tx <- plot_usmap(data = dat, 
                   regions = "counties", 
                   values = "d_marg_change", 
           include = c(fip_tx),
           labels = TRUE, 
           label.size = 0.1) +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "win margin") +
  theme_void() +
  theme(legend.position = "none") 

p_tx$layers[[2]]$aes_params$size <- 2.5
tx1 <- print(p_tx)

tx2 <- plot_usmap(data = dat, regions = "counties", values = "d_marg_change", 
           include = "Texas") +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "Win Margin") +
  theme_void()


tx_comp <- ggarrange(tx1, tx2, ncol=2)
texas_comp <- annotate_figure(tx_comp,
               top = text_grob("Texas 2016-2020: Change in Democratic \n Win Margins in Hispanic-Majority Counties", face = "bold", size = 13), 
               bottom = text_grob("Source: Pew",
                                  hjust = 1, x = 0.85, face = "italic", size = 10))

ggsave("~/Desktop/R studio/carine-h.github.io/images/10_texas_comp.png", texas_comp)



```

```{r}
#####-------------------------------------------------------------------------------#
##### New Mexico map ####
#####-------------------------------------------------------------------------------#
p_nm <- plot_usmap(data = dat, 
                   regions = "counties", 
                   values = "d_marg_change", 
           include = c(fip_nm),
           labels = TRUE) +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "win margin") +
  theme_void() +
  theme(legend.position = "none") 

p_nm$layers[[2]]$aes_params$size <- 4
nm1 <- print(p_nm)

nm2 <- plot_usmap(data = dat, regions = "counties", values = "d_marg_change", 
           include = "New Mexico") +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "Win Margin") +
  theme_void()


nm_comp <- ggarrange(nm1, nm2, ncol=2)
newmexico_comp <- annotate_figure(nm_comp,
               top = text_grob("New Mexico 2016-2020: Change in Democratic \n Win Margins in Hispanic-Majority Counties", face = "bold", size = 13), 
               bottom = text_grob("Source: Pew",
                                  hjust = 1, x = 0.85, face = "italic", size = 10))

ggsave("~/Desktop/R studio/carine-h.github.io/images/10_newmexico_comp.png", newmexico_comp)

```

```{r}
#####-------------------------------------------------------------------------------#
##### California map ####
#####-------------------------------------------------------------------------------#

p_ca <- plot_usmap(data = dat, 
                   regions = "counties", 
                   values = "d_marg_change", 
           include = fip_ca,
           labels = TRUE) +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "Win Margin") +
  theme_void() +
  theme(legend.position = "none") 

p_ca$layers[[2]]$aes_params$size <- 4
ca1 <- print(p_ca)

ca2 <- plot_usmap(data = dat, regions = "counties", values = "d_marg_change", 
           include = "California") +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "Win Margin") +
  theme_void()


ca_comp <- ggarrange(ca1, ca2, ncol=2)
cali_comp <- annotate_figure(ca_comp,
               top = text_grob("California 2016-2020: Change in Democratic \n Win Margins in Hispanic-Majority Counties", face = "bold", size = 13), 
               bottom = text_grob("Source: Pew",
                                  hjust = 1, x = 0.85, face = "italic", size = 10))

ggsave("~/Desktop/R studio/carine-h.github.io/images/10_cali_comp.png", cali_comp)

```
```{r}
#####-------------------------------------------------------------------------------#
##### Arizona map ####
#####-------------------------------------------------------------------------------#

p_az <- plot_usmap(data = dat, 
                   regions = "counties", 
                   values = "d_marg_change", 
           include = fip_az,
           labels = TRUE) +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "Win Margin") +
  theme_void() +
  theme(legend.position = "none") 

p_az$layers[[2]]$aes_params$size <- 4
az1 <- print(p_az)

az2 <- plot_usmap(data = dat, regions = "counties", values = "d_marg_change", 
           include = "Arizona") +
  scale_fill_gradient2(
    high = "blue", 
    mid = "white",
    low = "red", 
    breaks = c(-50,-25,0,25,50), 
    limits = c(-50,50),
    name = "Win Margin") +
  theme_void()


az_comp <- ggarrange(az1, az2, ncol=2)
arizona_comp <- annotate_figure(az_comp,
               top = text_grob("Arizona 2016-2020: Change in Democratic \n Win Margins in Hispanic-Majority Counties", face = "bold", size = 13), 
               bottom = text_grob("Source: Pew",
                                  hjust = 1, x = 0.85, face = "italic", size = 10))

ggsave("~/Desktop/R studio/carine-h.github.io/images/10_arizona_comp.png", arizona_comp)
```
Arizona, like Florida, had blue counties (that remained blue) lose part of their vote share to Trump in 2020. 

# Reason 1: Rhetoric 
- Trump's rhetoric resonated better among Hispanic voters, especially in Florida
```{r}
#####-------------------------------------------------------------------------------#
##### Trump vs. Biden campaign speeches with General Hispanic messaging ####
#####-------------------------------------------------------------------------------#

#### General overview of buzz words ####

## pre-process: make a `quanteda` corpus from dataframe
speech_corpus <- corpus(speech_df, text_field = "text", docid_field = "url")

## pre-process: tokenize, clean, select n-grams
speech_toks_hispanic <- tokens(speech_corpus, 
    remove_punct = TRUE,
    remove_symbols = TRUE,
    remove_numbers = TRUE,
    remove_url = TRUE) %>% 
    tokens_tolower() %>%
    tokens_remove(pattern=c("joe","biden","donald","trump","president","kamala","harris")) %>%
    tokens_remove(pattern=stopwords("en")) %>%
    tokens_select(min_nchar=3) %>%
    tokens_ngrams(n=2) %>%
    tokens_select(pattern = c("hisp*", "communism", "latin*", "socialism", "dict*", "cuba*", "maduro", 
                              "castro*"), padding = TRUE)

print(speech_toks_hispanic)

## pre-process: make doc-freq matrix
speech_dfm_hispanic <- dfm(speech_toks_hispanic, groups = "candidate")

## summarise and visualise
tstat_freq_hispanic <- textstat_frequency(speech_dfm_hispanic)
head(tstat_freq_hispanic, 100)

textplot_wordcloud(speech_dfm_hispanic, color = c("red", "blue"), comparison = T)

trump_keyness_hispanic <- textstat_keyness(speech_dfm_hispanic, target = "Donald Trump")
textplot_keyness(trump_keyness_hispanic, 
                 color = c("blue", "red"), 
                 n = 10) 


#####-------------------------------------------------------------------------------#
##### Trump vs. Biden campaign speeches on cuba/venezula/nicaragua ####
#####-------------------------------------------------------------------------------#


## pre-process: make a `quanteda` corpus from dataframe
speech_corpus <- corpus(speech_df, text_field = "text", docid_field = "url")

## pre-process: tokenize, clean, select n-grams
speech_toks_comm <- tokens(speech_corpus, 
    remove_punct = TRUE,
    remove_symbols = TRUE,
    remove_numbers = TRUE,
    remove_url = TRUE) %>% 
    tokens_tolower() %>%
    tokens_remove(pattern=c("joe","biden","donald","trump","president","kamala","harris")) %>%
    tokens_remove(pattern=stopwords("en")) %>%
    tokens_select(min_nchar=3) %>%
    tokens_ngrams(n=2) %>%
    tokens_select(pattern = c("maduro*", "castro*", "cuba*", "venezuel*", "nicarag*"))

print(speech_toks_comm)

## pre-process: make doc-freq matrix
speech_dfm_comm <- dfm(speech_toks_comm, groups = "candidate")

## summarise and visualise
tstat_freq_comm <- textstat_frequency(speech_dfm_comm)
head(tstat_freq_comm, 100)

textplot_wordcloud(speech_dfm_comm, color = c("red", "blue"), comparison = T)

trump_keyness_comm <- textstat_keyness(speech_dfm_comm, target = "Donald Trump")
textplot_keyness(trump_keyness_comm, 
                 color = c("red", "blue"), 
                 margin = 0.4) 

# this was too broad I think 
# c("communis*", "sociali*","cuba*", "maduro", 
#                              "castro*", "venezuela*", "chavez", "chavi*"), padding = TRUE)



#####-------------------------------------------------------------------------------#
##### Trump vs. Biden campaign speeches on socialism/communism ####
#####-------------------------------------------------------------------------------#

speech_corpus <- corpus(speech_df, text_field = "text", docid_field = "url")

## pre-process: tokenize, clean, select n-grams
speech_toks_soc <- tokens(speech_corpus, 
    remove_punct = TRUE,
    remove_symbols = TRUE,
    remove_numbers = TRUE,
    remove_url = TRUE) %>% 
    tokens_tolower() %>%
    tokens_remove(pattern=c("joe","biden","donald","trump","president","kamala","harris")) %>%
    tokens_remove(pattern=stopwords("en")) %>%
    tokens_select(min_nchar=3) %>%
    tokens_ngrams(n=2) %>%
    tokens_select(pattern = c("sociali*", "communis*"))

print(speech_toks_soc)

## pre-process: make doc-freq matrix
speech_dfm_soc <- dfm(speech_toks_soc, groups = "candidate")

## summarise and visualise
tstat_freq_soc  <- textstat_frequency(speech_dfm_soc)
head(tstat_freq_soc, 100)

textplot_wordcloud(speech_dfm_comm, color = c("red", "blue"), comparison = T)

trump_keyness_soc <- textstat_keyness(speech_dfm_soc, target = "Donald Trump")
textplot_keyness(trump_keyness_soc, 
                 color = c("red", "blue"), 
                 margin = 0.4) 
```

# Reason 2: Ad Spending 
```{r}
# cleaning and joining FB data 
fb_loc <- fb_location_df %>%
  rename("state" = "Location Name", 
         "spent" = "Amount Spent (USD)") %>%
  mutate(state = state.abb[match(state,state.name)]) %>%
  na.omit() 
 fb_loc <- fb_loc %>%
   mutate(ad_spent = as.numeric(fb_loc$spent))
 
# plotting amount spent per state on FB ads
plot_usmap(data = fb_loc, regions = "state", values = "ad_spent") +
  scale_fill_gradient2(
    high = "indianred", 
    mid = "white",
    low = "grey", 
    name = "Ad Money Spent")+ 
    theme_void()+
  labs(title = "2020: Total Facebook Ad Money Spent", 
       subtitle = "Presidential Campaigns and their Supporters", 
       caption = "Data: Facebook") + 
  theme(plot.title = element_text(face = "bold"))
ggsave("~/Desktop/R studio/carine-h.github.io/images/10_fb_spending.png", height = 5, width =10)
```



  
