---
title: "blog_2_rmd"
author: "Carine Hajjar"
date: "9/20/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
####----------------------------------------------------------#
#### Pre-amble ####
####----------------------------------------------------------#

## install via `install.packages("name")`
library(tidyverse)
library(ggplot2)
library(jtools) # for plot_summ 
library(huxtable)
library(gt)
library(gtsummary)
library(broom.mixed)
## set working directory here
setwd("~")

```


```{r}
####----------------------------------------------------------#
#### The relationship between economy and PV ####
####----------------------------------------------------------#

# reading in the data
economy_df <- read_csv("~/Desktop/R studio/carine-h.github.io/data/econ.csv") 
popvote_df <- read_csv("~/Desktop/R studio/carine-h.github.io/data/popvote_1948-2016 copy.csv") 

# I updated my_pretty_theme 
my_pretty_theme <- theme_bw() + 
  theme(panel.border = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, face = "italic"),
        plot.title   = element_text(size = 15, hjust = 0.5, face = "bold"), 
        axis.text    = element_text(size = 12),
        strip.text   = element_text(size = 18),
        axis.line    = element_line(colour = "black"),
        legend.text = element_text(size = 12))

```

# BLOG MATERIAL
*Extension: Heeterogenous Predictive Power of the Economy*
Is the effect of the economy stronger for sitting presidents running for re-election compared to same-party heirs seeking the office?
Does the predictive power of economy change across time? If so, why? 

```{r}
# joining popvote and econ data and using second quarter of the election year, per Achen and Bartels reading 
data_1 <- popvote_df %>% 
  select(year, winner, pv2p, incumbent, incumbent_party) %>%
  left_join(economy_df)%>%
  filter(quarter == 2)

# GDP yearly vs. GDP quarter 
fit_1 <- lm(pv2p ~ GDP_growth_qt, data = data_1)
summary(fit_1)

```
This model tells us that incumbent winners see are likely associated with a GDP growth in their election year. However, GDP growth in the quarter concurrent with their reelection is actually a stronger predictor, consistent with Achen and Bartels' assessment of economic  voting. 

Above, I use GDP growth to model incumbent wins. Here, more consistent with Achen and Bartel's study, I'll use "quarterly growth rates in real income" (which they believe to be a better indication of electoral preference in the popular vote share than GDP growth) in the short-term and long-term of a presidential election.  
```{r}
data_year_growth <- popvote_df %>% 
  select(year, winner, pv2p, incumbent, incumbent_party) %>%
  left_join(economy_df)%>%
# this column will tell us whether or not the winner was an incumbent or not  
  mutate(incumbent_winner = ifelse(winner == TRUE & incumbent == TRUE, TRUE, FALSE))%>%
  select(year, winner, pv2p, winner, incumbent, incumbent_party, incumbent_winner, RDI_growth, GDP, GDP_growth_yr, GDP_growth_qt, inflation, unemployment)

summary(lm(winner ~ RDI_growth, data = data_year_growth))
# is winner best way? 
# maybe PV2? 
```

```{r}
# FIT 3: Second-Quarter Incumbent PV2P and RDI Growth
data_quarter_growth <- popvote_df %>% 
  select(year, winner, pv2p, incumbent_party) %>%
  left_join(economy_df)%>%
  filter(quarter == 2, 
         incumbent_party == TRUE) %>%
  select(year, winner, pv2p, winner, incumbent_party, RDI_growth, GDP, GDP_growth_yr, GDP_growth_qt, inflation, unemployment)

lm(pv2p ~ RDI_growth, data = data_quarter_growth)
summary(data_quarter_growth)

# this graphs the incumbent's popular vote according to second quarter real income growth and includes a line for standard deviation 
  ggplot(data = data_quarter_growth, aes(x=RDI_growth, y=pv2p,
             label=year)) + 
  geom_text() +
# inserts error 
  geom_smooth(method="lm", formula = y ~ x) +
  geom_hline(yintercept=50, lty=2) +
  geom_vline(xintercept=0.01, lty=2) + # median
  xlab("Second Quarter RDI Growth") +
  ylab("Popular Two Party Vote Share") +
  labs(title = "RDI Growth and Vote Shares for Incumbent Parties", 
       subtitle = "Second Quarter of Election Year", 
       caption = "Data from 1948-2016 U.S. elections and economies") +
  my_pretty_theme

ggsave("~/Desktop/R studio/carine-h.github.io/images/RDI_pv2p.png", height = 2, width = 4)

## fit a model
lm_econ <- lm(pv2p ~ GDP_growth_qt, data = data_quarter_growth)
summary(lm_econ)

ggplot(data = data_quarter_growth, aes(x=GDP_growth_qt, y=pv2p,
             label=year)) + 
  geom_text() +
# inserts error 
  geom_smooth(method="lm", formula = y ~ x) +
  geom_hline(yintercept=50, lty=2) +
  geom_vline(xintercept=0.01, lty=2) + # median
  xlab("Second Quarter GDP Growth") +
  ylab("Popular Two Party Vote Share") +
  labs(title = "GDP Growth and Vote Shares for Incumbent Parties", 
       subtitle = "Second Quarter of Election Year", 
       caption = "Data from 1948-2016 U.S. elections and economies") +
  my_pretty_theme

ggsave("~/Desktop/R studio/carine-h.github.io/images/GDP_pv2p.png", height = 2, width = 4)
```


Answering: Is the effect of the economy (using 2nd quarter GDP growth, RDI, Unemployment) stronger for sitting presidents running for re-election compared to same-party heirs seeking the office?

# Incumbent Non-Incumbent Economic Regressions
```{r, results = 'asis'}
##########
# INCUMBENT/NON-INCUMBENT MODELS
##########

# DATA FOR INCUMBENTS
data_incumbent <- popvote_df %>% 
  select(year, winner, pv2p, incumbent_party, incumbent) %>%
  left_join(economy_df)%>%
  filter(quarter == 2, 
         incumbent == TRUE) %>%
  select(year, winner, pv2p, winner, incumbent_party, RDI_growth, GDP, GDP_growth_yr, GDP_growth_qt, inflation, unemployment)

# impact of economy for INCUMBENTS 
lm_incumbent <- lm(pv2p ~ GDP_growth_qt + RDI_growth, data = data_incumbent)

# DATA FOR NON-INCUMBENTS
data_nonincumbent <- popvote_df %>% 
  select(year, winner, pv2p, incumbent_party, incumbent) %>%
  left_join(economy_df)%>%
  filter(quarter == 2, 
         incumbent == FALSE,
         incumbent_party == TRUE) %>%
  select(year, winner, pv2p, winner, incumbent_party, RDI_growth, GDP, GDP_growth_yr, GDP_growth_qt, inflation, unemployment)

# impact of economy for NON-INCUMBENTS
lm_nonincumbent <- lm(pv2p ~ GDP_growth_qt + RDI_growth, data = data_nonincumbent)


plot_summs(lm_incumbent, lm_nonincumbent, coefs = c( "2nd Quarter \n GDP Growth" = "GDP_growth_qt", "2nd Quarter \n RDI Growth" = "RDI_growth"), scale = TRUE, robust = TRUE, legend.title = "Incumbent v. \n Non-incumbent Candidate")+
  my_pretty_theme+
  labs(title = "Impact of Economic Performance on the \n Two-Party Vote Share",
       subtitle = "Second Quarter GDP and Real Income Growth (1948-2016)", 
       caption = "Data from 1948-2016 U.S. elections and economies", 
       y = "Economic Metric", 
       x = "Estimate")
ggsave("~/Desktop/R studio/carine-h.github.io/images/GDP_RDI_model.png", height = 2, width = 4)


###############
# UNEMPLOYMENT 
###############
# Unemployment
lm_UE_inc <-  lm(pv2p ~ unemployment, data = data_incumbent)
lm_UE_noninc <-  lm(pv2p ~ unemployment, data = data_nonincumbent)

plot_summs(lm_UE_inc, lm_UE_noninc, coefs = c("Unemployment" = "unemployment"), scale = TRUE, robust = TRUE, legend.title = "Incumbent v. \n Non-incumbent Candidate")+
    my_pretty_theme+
  labs(title = "Impact of Unemployment on the \n Two-Party Vote Share",
       subtitle = "Second Quarter Unemployment (1948-2016)",
       caption = "Data from 1948-2016 U.S. elections and economies", 
       y = "", 
       x = "Estimate")
ggsave("~/Desktop/R studio/carine-h.github.io/images/UE_model.png", height = 2, width = 4)

# Regression summaries 
export_summs(lm_incumbent, lm_nonincumbent)
export_summs(lm_UE_inc, lm_UE_noninc)

```



# ECON OVERTIME 
Does the predictive power of economy change across time? If so, why? 

Going back to the incumbent reelection and its relationship with the economy, it's important to consider the correlation of these variables overtime. When is the economy the most important? 

COMMENTARY

This is a crucial question at the moment given the pandemic's effect on the American economy. Will the economy be a bigger predictor now more than ever?
```{r}
data_incumbent_80 <- popvote_df %>% 
  select(year, winner, pv2p, incumbent_party, incumbent) %>%
  left_join(economy_df)%>%
  filter(quarter == 2, 
         year <= 1980) %>%
  select(year, winner, pv2p, winner, incumbent_party, RDI_growth, GDP, GDP_growth_yr, GDP_growth_qt, inflation, unemployment)

data_incumbent_after80 <- popvote_df %>% 
  select(year, winner, pv2p, incumbent_party, incumbent) %>%
  left_join(economy_df)%>%
  filter(quarter == 2, 
         year > 1980) %>%
  select(year, winner, pv2p, winner, incumbent_party, RDI_growth, GDP, GDP_growth_yr, GDP_growth_qt, inflation, unemployment)

lm_incumbent_80 <- lm(pv2p ~ GDP_growth_qt, data = data_incumbent_80)
summary(lm_incumbent_80)

lm_incumbent_after80 <- lm(pv2p ~ GDP_growth_qt, data = data_incumbent_after80)
summary(lm_incumbent_after80)


export_summs(lm_incumbent_80, lm_incumbent_after80)
```


```{r}
data_inc <- popvote_df %>% 
  select(year, winner, pv2p, incumbent_party, incumbent) %>%
  left_join(economy_df)%>%
  filter(quarter == 2, 
         incumbent == TRUE) %>%
  select(year, winner, pv2p, winner, incumbent_party, RDI_growth, GDP, GDP_growth_yr, GDP_growth_qt, inflation, unemployment)%>%
  rename(Unemployment = unemployment)

fit_econ <- lm(pv2p ~ Unemployment, data = data_inc)

tbl_regression(fit_econ, intercept = TRUE) %>%
  as_gt() %>%
  tab_header(title = md("**Relationship Between the Vote Share and GDP Growth and Unemployment \n for Incumbent Presidential Campaigns**"), 
              subtitle = md("*Second Quarter, 1948-2016*"))%>%
  tab_source_note(md("**Source:** 1948-2016 election results and economic performance")) %>% 
  tab_footnote( 
    footnote = md("Calculated via a permutation test"),
    locations = cells_column_labels(columns = vars(p.value))
  )


```







